\newpage
\centerline{\Large\bf Preface}
\par
\medskip
\par
Please note that this is an alpha release of the package,
probably a 0.5 release in anticipation of a 1.0 version that
will be useful and self-contained.
\par
The serial code and the threaded code using the Solaris thread package
is relatively stable, meaning that every once in a while we shake
out a bug but the code does factor and solve sparse linear systems
relatively efficiently and in a fairly easy to use manner.
The MPI version of the code is still under development, and we are
reluctant to release the present code because we want the present code 
base for the three versions --- serial, threaded and MPI --- to be
coherent. We don't anticipate any major changes to the code base as
a result of the MPI development, but we offer no guarantees.
\par
The {\bf SPOOLES} library depends quite strongly on the objects
found in the {\bf SMOOTH 1.1} library for ordering sparse matrices.
The code conventions are the same between the two libraries.
At some point in the future we will merge the two into one library.
\par
This reference document is not complete.
All the objects are documented, their data structures and
methods completely so.
The background on the objects, the ``why'' and ``how'' of the
design is lacking.
The reader simply interested using the library should have little
difficulty.
The reader wanting to know more will appreciate a later version of
this document that will contain a much better background discussion 
on motivation and software and will include a respectable
bibliography.
\par
Please communicate all questions and comments to
{\tt cleve.ashcraft@boeing.com}.
