\chapter{The Serial Bridge Object and Driver}
\label{chapter:serial}
\par
\section{The \texttt{Bridge} Data Structure}
\label{section:Bridge:dataStructure}
\par
The {\tt Bridge} structure has the following fields.
\begin{itemize}
%
\item
{\tt int prbtype} : problem type
\begin{itemize}
\item {\tt 1} --- vibration, a multiply with $B$ is required.
\item {\tt 2} --- buckling, a multiply with $A$ is required.
\item {\tt 3} --- simple, no multiply is required.
\end{itemize}
\item
{\tt int neqns} : number of equations, 
i.e., number of vertices in the graph.
\item
{\tt int mxbsz} : block size for the Lanczos process.
\item
{\tt int seed} : random number seed used in the ordering.
\item
{\tt InpMtx *A} : matrix object for $A$
\item
{\tt InpMtx *B} : matrix object for $B$
\item
{\tt Pencil *pencil} : object to hold linear combination of $A$ and $B$.
\item
{\tt ETree *frontETree} : object that defines the factorizations,
e.g., the number of fronts, the tree they form, the number of
internal and external rows for each front, and the map from
vertices to the front where it is contained.
\item
{\tt IVL *symbfacIVL} : object that contains the symbolic
factorization of the matrix.
\item
{\tt SubMtxManager *mtxmanager} : object that manages the
\texttt{SubMtx} objects that store the factor entries and are used
in the solves.
\item
{\tt FrontMtx *frontmtx} : object that stores the $L$, $D$ and $U$
factor matrices.
\item
{\tt IV *oldToNewIV} : object that stores old-to-new permutation vector.
\item
{\tt IV *newToOldIV} : object that stores new-to-old permutation vector.
\item
{\tt DenseMtx *X} : dense matrix object that is used during the matrix
multiples and solves.
\item
{\tt DenseMtx *Y} : dense matrix object that is used during the matrix
multiples and solves.
\item
{\tt int msglvl} : message level for output.
When 0, no output, When 1, just statistics and cpu times.
When greater than 1, more and more output.
\item
{\tt FILE *msgFile} : message file for output.
When \texttt{msglvl} $>$ 0, \texttt{msgFile} must not be \texttt{NULL}.
\end{itemize}
\par
\section{Prototypes and descriptions of \texttt{Bridge} methods}
\label{section:Bridge:proto}
\par
This section contains brief descriptions including prototypes
of all methods that belong to the {\tt Bridge} object.
\par
%=======================================================================
\begin{enumerate}
%-----------------------------------------------------------------------
\item
\begin{verbatim}
int Setup ( void *data, int *pprbtype, int *pneqns, int *pmxbsz,
            InpMtx *A, InpMtx *B, int *pseed, int *pmsglvl, FILE *msgFile ) ;
\end{verbatim}
\index{Setup@{\tt Setup()}}
\noindent All calling sequence parameters are pointers to more
easily allow an interface with Fortran.
\begin{itemize}
\item {\tt void *data} --- a pointer to the {\tt Bridge} object.
\item {\tt int *pprbtype} --- {\tt *pprbtype} holds the problem type.
   \begin{itemize}
   \item {\tt 1} --- vibration, a multiply with $B$ is required.
   \item {\tt 2} --- buckling, a multiply with $A$ is required.
   \item {\tt 3} --- simple, no multiply is required.
   \end{itemize}
\item {\tt int *pneqns} --- {\tt *pneqns} is the number of equations.
\item {\tt int *pmxbsz} --- {\tt *pmxbsz} is an upper bound on the
block size.
\item {\tt InpMtx *A} --- {\tt A} is a {\bf SPOOLES} object that
holds the matrix $A$.
\item {\tt InpMtx *B} --- {\tt B} is a {\bf SPOOLES} object that
holds the matrix $B$. For an ordinary eigenproblem, $B$ is the
identity and {\tt B} is {\tt NULL}.
\item {\tt int *pseed} --- {\tt *pseed} is a random number seed.
\item {\tt int *pmsglvl} --- {\tt *pmsglvl} is a message level for
the bridge methods and the {\bf SPOOLES} methods they call.
\item {\tt FILE *pmsglvl} --- {\tt msgFile} is the message file
for the bridge methods and the {\bf SPOOLES} methods they call.
\end{itemize}
This method must be called in the driver program prior to invoking
the eigensolver via a call to {\tt lanczos\_run()}.
It then follows this sequence of action.
\begin{itemize}
\item
The method begins by checking all the input data,
and setting the appropriate fields of the {\tt Bridge} object.
\item
The {\tt pencil} object is initialized with {\tt A} and {\tt B}.
\item
{\tt A} and {\tt B} are converted to storage by rows and vector mode.
\item
A {\tt Graph} object is created that contains the sparsity pattern of
the union of {\tt A} and {\tt B}.
\item
The graph is ordered by first finding a recursive dissection partition,
and then evaluating the orderings produced by nested dissection and
multisection, and choosing the better of the two.
The {\tt frontETree} object is produced and placed into the {\tt bridge}
object.
\item
Old-to-new and new-to-old permutations are extracted from the front tree
and loaded into the {\tt Bridge} object.
\item
The vertices in the front tree are permuted, as well as the entries in 
{\tt A} and {\tt B}.
Entries in the lower triangle of {\tt A} and {\tt B} are mapped into the
upper triangle, and the storage modes of {\tt A} and {\tt B} are changed
to chevrons and vectors, in preparation for the first factorization.
\item
The symbolic factorization is then computed and loaded in the {\tt
Bridge} object.
\item
A {\tt FrontMtx} object is created to hold the factorization
and loaded into the {\tt Bridge} object.
\item
A {\tt SubMtxManager} object is created to hold the factor's
submatrices and loaded into the {\tt Bridge} object.
\item
Two {\tt DenseMtx} objects are created to be used during the matrix
multiplies and solves.
\end{itemize}
The {\tt A} and {\tt B} matrices are now in their permuted ordering,
i.e., $PAP^T$ and $PBP^T$, and all data structures are with respect to
this ordering. After the Lanczos run completes, any generated
eigenvectors must be permuted back into their original ordering using
the {\tt oldToNewIV} and {\tt newToOldIV} objects.
\par \noindent {\it Return value:}
\begin{center}
\begin{tabular}[t]{rl}
~1 & normal return \\
-1 & \texttt{data} is \texttt{NULL} \\
-2 & \texttt{pprbtype} is \texttt{NULL} \\
-3 & \texttt{*pprbtype} is invalid \\
-4 & \texttt{pneqns} is \texttt{NULL} \\
-5 & \texttt{*pneqns} is invalid
\end{tabular}
\begin{tabular}[t]{rl}
-6 & \texttt{pmxbsz} is \texttt{NULL} \\
-7 & \texttt{*pmxbsz} is invalid \\
-8 & \texttt{A} and \texttt{B} are \texttt{NULL} \\
-9 & \texttt{seed} is \texttt{NULL} \\
-10 & \texttt{msglvl} is \texttt{NULL} \\
-11 & $\texttt{msglvl} > 0$ and \texttt{msgFile} is \texttt{NULL}
\end{tabular}
\end{center}
%-----------------------------------------------------------------------
\item
\begin{verbatim}
void Factor ( double *psigma, double *ppvttol, void *data, int *pinertia, int *perror ) ;
\end{verbatim}
\index{Factor@{\tt Factor()}}
This method computes the factorization of $A - \sigma B$.
All calling sequence parameters are pointers to more
easily allow an interface with Fortran.
\begin{itemize}
\item {\tt double *psigma} --- the shift parameter $\sigma$ is
      found in {\tt *psigma}.
\item {\tt double *ppvttol} --- the pivot tolerance is
      found in {\tt *ppvttol}. When ${\tt *ppvttol} = 0.0$, 
      the factorization is computed without pivoting for stability.
      When ${\tt *ppvttol} > 0.0$, the factorization is computed 
      with pivoting for stability, and all offdiagonal entries 
      have magnitudes bounded above by $1/({\tt *ppvttol})$.
\item {\tt void *data} --- a pointer to the {\tt Bridge} object.
\item {\tt int *pinertia} --- on return, {\tt *pinertia} holds the 
      number of negative eigenvalues.
\item {\tt int *perror} --- on return, {\tt *perror} holds an
      error code.
      \begin{center}
      \begin{tabular}[t]{rl}
      ~1 & error in the factorization \\
      ~0 & normal return \\
      -1 & \texttt{psigma} is \texttt{NULL}
      \end{tabular}
      \begin{tabular}[t]{rl}
      -2 & \texttt{ppvttol} is \texttt{NULL} \\
      -3 & \texttt{data} is \texttt{NULL} \\
      -4 & \texttt{pinertia} is \texttt{NULL}
      \end{tabular}
      \end{center}
\end{itemize}
%-----------------------------------------------------------------------
\item
\begin{verbatim}
void MatMul ( int *pnrows, int *pncols, double X[], double Y[],
              int *pprbtype, void *data ) ;
\end{verbatim}
\index{MatMul@{\tt MatMul()}}
This method computes a multiply of the form $Y = I X$, $Y = A X$
or $Y = B X$.
All calling sequence parameters are pointers to more
easily allow an interface with Fortran.
\begin{itemize}
\item {\tt int *pnrows} --- {\tt *pnrows} contains the number of
      rows in $X$ and $Y$.
\item {\tt int *pncols} --- {\tt *pncols} contains the number of
      columns in $X$ and $Y$.
\item {\tt double X[]} --- this is the $X$ matrix, stored column
      major with leading dimension {\tt *pnrows}.
\item {\tt double Y[]} --- this is the $Y$ matrix, stored column
      major with leading dimension {\tt *pnrows}.
\item {\tt int *pprbtype} --- {\tt *pprbtype} holds the problem type.
   \begin{itemize}
   \item {\tt 1} --- vibration, a multiply with $B$ is required.
   \item {\tt 2} --- buckling, a multiply with $A$ is required.
   \item {\tt 3} --- simple, no multiply is required.
   \end{itemize}
\item {\tt void *data} --- a pointer to the {\tt Bridge} object.
\end{itemize}
%-----------------------------------------------------------------------
\item
\begin{verbatim}
void Solve ( int *pnrows, int *pncols, double X[], double Y[],
             void *data, int *perror ) ;
\end{verbatim}
\index{Solve@{\tt Solve()}}
This method solves $(A - \sigma B) X = Y$, where 
$(A - \sigma B)$ has been factored by a previous call to {\tt Factor()}.
All calling sequence parameters are pointers to more
easily allow an interface with Fortran.
\begin{itemize}
\item {\tt int *pnrows} --- {\tt *pnrows} contains the number of
      rows in $X$ and $Y$.
\item {\tt int *pncols} --- {\tt *pncols} contains the number of
      columns in $X$ and $Y$.
\item {\tt double X[]} --- this is the $X$ matrix, stored column
      major with leading dimension {\tt *pnrows}.
\item {\tt double Y[]} --- this is the $Y$ matrix, stored column
      major with leading dimension {\tt *pnrows}.
\item {\tt void *data} --- a pointer to the {\tt Bridge} object.
\item {\tt int *perror} --- on return, {\tt *perror} holds an
      error code.
      \begin{center}
      \begin{tabular}[t]{rl}
      ~1 & normal return \\
      -1 & \texttt{pnrows} is \texttt{NULL} \\
      -2 & \texttt{pncols} is \texttt{NULL}
      \end{tabular}
      \begin{tabular}[t]{rl}
      -3 & \texttt{X} is \texttt{NULL} \\
      -4 & \texttt{Y} is \texttt{NULL} \\
      -5 & \texttt{data} is \texttt{NULL}
      \end{tabular}
      \end{center}
      \end{itemize}
%-----------------------------------------------------------------------
\item
\begin{verbatim}
int Cleanup ( void *data ) ;
\end{verbatim}
\index{Cleanup@{\tt Cleanup()}}
This method releases all the storage used by the {\bf SPOOLES}
library functions.
\par \noindent {\it Return value:}
1 for a normal return,
-1 if a {\tt data} is {\tt NULL}.
%-----------------------------------------------------------------------
\end{enumerate}
\par
\section{The \texttt{testSerial} Driver Program}
\label{section:Bridge:driver}
\par
A complete listing of the serial driver program is
found in chapter~\ref{chapter:serial_driver}.
The program is invoked by this command sequence.
\begin{verbatim}
testSerial msglvl msgFile parmFile seed inFileA inFileB
\end{verbatim}
where
\begin{itemize}
\item 
{\tt msglvl} is the message level for the {\tt Bridge}
methods and the {\bf SPOOLES} software.
\item 
{\tt msgFile} is the message file for the {\tt Bridge}
methods and the {\bf SPOOLES} software.
\item 
{\tt parmFile} is the input file for the parameters of the
eigensystem to be solved.
\item 
{\tt seed} is a random number seed
used by the {\bf SPOOLES} software.
\item 
{\tt inFileA} is the Harwell-Boeing file for the matrix $A$.
\item 
{\tt inFileB} is the Harwell-Boeing file for the matrix $B$.
\end{itemize}
This program is executed for some sample matrices by the
{\tt do\_ST\_*} shell scripts in the {\tt drivers} directory.
\par
Here is a short description of the steps in the driver program.
See Chapter~\ref{chapter:serial_driver} for the listing.
\begin{enumerate}
\item
The command line inputs are decoded.
\item
The header of the Harwell-Boeing file for $A$ is read.
This yields the number of equations.
\item
The parameters that define the eigensystem to be solved
are read in from the {\tt parmFile} file.
\item
The Lanczos eigensolver workspace is initialized.
\item
The Lanczos communication structure is filled with some parameters.
\item
The $A$ and possibly $B$ matrices are read in from the
Harwell-Boeing files and converted into {\tt InpMtx} objects
from the {\bf SPOOLES} library.
\item
The linear solver environment is then initialized via a call to 
{\tt Setup()}.
\item
The eigensolver is invoked via a call to {\tt lanczos\_run()}.
The {\tt FactorMT()}, {\tt SolveMT()} and {\tt MatMulMT()}
methods are passed to this routine.
\item
The eigenvalues are extracted and printed via a call to
{\tt lanczos\_eigenvalues()}.
\item
The eigenvectors are extracted and printed via calls to
{\tt lanczos\_eigenvector()}.
\item
The eigensolver working storage is free'd via a call to
{\tt lanczos\_free()}.
\item
The linear solver working storage is free'd via a call to
{\tt Cleanup()}.
\end{enumerate}
