\par
\section{Driver programs for the {\tt Drand} object}
\label{section:Drand:drivers}
\par
This section contains brief descriptions of the driver programs.
\par
%=======================================================================
\begin{enumerate}
%-----------------------------------------------------------------------
\item
\begin{verbatim}
testDrand msglvl msgFile distribution param1 param2 seed1 seed2 n
\end{verbatim}
This driver program test the {\tt Drand} random number generator.
\par
\begin{itemize}
\item
The {\tt msglvl} parameter determines the amount of output.
\item
The {\tt msgFile} parameter determines the message file --- if {\tt
msgFile} is {\tt stdout}, then the message file is {\it stdout},
otherwise a file is opened with {\it append} status to receive any
output data.
\item
The {\tt distribution} parameter specifies the mode of the object.
If {\tt 1}, the distribution is uniform.
If {\tt 2}, the distribution is normal.
\item
When {\tt distribution = 1}, {\tt param1} is the lower bound for
the interval.
When {\tt distribution = 2}, {\tt param1} is the mean for the
normal distribution.
\item
When {\tt distribution = 1}, {\tt param2} is the upper bound for
the interval.
When {\tt distribution = 2}, {\tt param2} is the variance for the
normal distribution.
\item
{\tt seed1} is the first random number seed.
\item
{\tt seed2} is the second random number seed.
\item
{\tt n} is the length of the vector of random numbers to be
generated.
\end{itemize}
%-----------------------------------------------------------------------
\end{enumerate}
